Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-25 paper/paper.tex
conclusions

\tol's structural properties seem to be similar to observed dwarf
elliptical galaxies. 
The sample of 7 dwarf ellipticals (dE) observed by
\cite{2003AJ....126.1794G} show similar properties as \tol. 
For instance, the dE IC 3344 and IC 794 have similar absolute $M_V$
magnitudes (-17.10 and -17.19) and sizes (0.90 kpc and 0.82 kpc) than
\tol. These galaxies have velocity dispersions of $31.1$ \kms and
$44.6$ \kms, respectively,  and their rotational velocities are lower
than $1$\kms.  
This coincidence supports the idea that the multiphase model is
the correct one as it provides at least some consistency with
kinematic properties of other dwarf elliptical galaxies.

Using the constraints for the velocity dispersion $\sigma$  in a
spherical system localized in a region of size $r$ we can estimate a
value or the dynamical mass within $r$:

\begin{equation}
M_{\rm dyn} = 3 \frac{\sigma ^{2}r}{G} = 3.48\times10^{9}
\left(\frac{\sigma}{100\ \mathrm{km\ s}^{-1}}\right)^2\left(\frac{r}{\mathrm{kpc}}\right)
M_{\odot}, 
\end{equation}

Assuming that the \lya\ emission is entirely powered by star formation 
we use the 3D half-light radius $r_s=2.25$ kpc as the typical size
for the HI region, in the multiphase model the clump velocity dispersion
$\sigma_{\rm{cl}}=$\sigmaclump corresponds to a dynamical mass of
$M_{\rm dyn}=2.31\pm0.04 \times 10^{9}$ $M_{\odot}$.

Following the results of \cite{2011ApJ...726..108T}, who made a
detailed analysis for dwarf ellipticals, we have that such dynamical
mass can be explained if \tol\ is hosted by a dark matter halo of at
least $10^{11}$ M$_{\odot}$ in mass. 
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the \emph{median} mass-concentration
relationship found in the Bolshoi simulation \citep{2012MNRAS.423.3018P}.

This results would imply that \tol\ is part of the dwarf ellipticals
class. However, there are two problems that remain. First, dwarf
ellipticals are found either in clusters or groups of galaxies; while
\tol\ does not belong either to a group or a cluster
\citep{2001AJ....121..169F,}. The second problem is that, unlike dwarf
ellipticals, \tol\ hosts bright an compact startburst; imaging and
spectroscopy studies suggest that the startburst started less than 4
Myr ago, while the low surface brightness regins are consistent with
an age below one Gyr \citep{2001AJ....121..169F}.

New observational tests are needed to clarify the kinematic nature of
\tol. 
We suggest that integral field unit measurements 
resolving its spatial extent are up to the task. 
\tol\ spans a region of $4$ arcseconds,
an instrument such as the Multi Unit Spectroscopic Explorer (MUSE)
\citep{2014Msngr.157...13B} with its
nominal $0.2$ arcseconds spatial sampling over a $1.0$ arcminute field
in wide-field mode could provide a coarse mapping of different
ionization lines to infer a kinematic map in the H$\alpha$ line,
following the kind of work presented by \citet{Herenz16} done with the Potsdam
Multi-Aperture Spectrometer (PMAS) \citet{PMAS} on nearby
\lya\ emitting galaxies.

Another observational test includes the measurement of the
\lya\ ionizing continuum escape fraction.
In the rotational model this fraction should be zero, while
the multiphase model predicts that averaging over all sightlines
it should be around $0.5^{+1.0}_{-0.4}$\%, with the possibility of strong
variations depending on viewing angle (i.e. some sightlines can have
an escape fraction close to $100\%$)\citep{Gronke2016} .

@@

It is the first time that an observed \lya\ profile can be fully
reproduced by either of these two kinematic conditions.
Furthermore, the kinematics of both models suggest a dynamical mass
at least ten times larger than the baryonic mass infered from observations.
We suggest that additional integral field unit observations are the
only way to be certain about the detailed kinematic structure in \tol.

@@

We acknowledge that a final confirmation of our kinematic findings has
to be provided by new observational tests of the kind we suggest.
If that were the case, this would open up a new line of more detailed
enquiries to understand how a compact dwarf galaxy can have such a
large dynamical mass.

>>>>

\tol's structural properties seem to be similar to observed dwarf
elliptical galaxies. 
The sample of 7 dwarf ellipticals (dE) observed by
\cite{2003AJ....126.1794G} show similar properties as \tol. 
For instance, the dE IC 3344 and IC 794 have similar absolute $M_V$
magnitudes (-17.10 and -17.19) and sizes (0.90 kpc and 0.82 kpc) than
\tol. These galaxies have velocity dispersions of $31.1$ \kms and
$44.6$ \kms, respectively,  and their rotational velocities are lower
than $1$ \kms.  
This coincidence leads us to prefer the multiphase model as it
provides at least some consistency with kinematic properties of other
dwarf elliptical galaxies.

Following the results of \cite{2011ApJ...726..108T}, who made a
detailed analysis for dwarf ellipticals, we can estimate a
value or the dynamical mass within $r$ using the constraints for the
velocity dispersion $\sigma$  in a spherical system localized in a
region of size $r$

\begin{equation}
M_{\rm dyn} = 3 \frac{\sigma ^{2}r}{G} = 3.48\times10^{9}
\left(\frac{\sigma}{100\ \mathrm{km\ s}^{-1}}\right)^2\left(\frac{r}{\mathrm{kpc}}\right)
M_{\odot}, 
\end{equation}
%
Assuming that the \lya\ emission is entirely powered by star formation 
we use the 3D half-light radius $r_s=2.25$ kpc as the typical size
for the HI region, in the multiphase model the clump velocity dispersion
$\sigma_{\rm{cl}}=$\sigmaclump corresponds to a dynamical mass of
$M_{\rm dyn}=2.31\pm0.04 \times 10^{9}$ $M_{\odot}$.

\cite{2011ApJ...726..108T} showed that  that such dynamical
mass can be explained if \tol\ is hosted by a dark matter halo of at
least $10^{11}$ M$_{\odot}$ in mass. 
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the \emph{median} mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P}.

These results would imply that \tol\ is part of the dwarf ellipticals
class. However, there are two puzzles that are opened up by this. 
First, dwarf ellipticals are found either in clusters or groups of galaxies; while
\tol\ does not belong either to a group or a cluster
\citep{2001AJ....121..169F,}. 
The second problem is that, unlike dwarf
ellipticals, \tol\ hosts bright an compact startburst; imaging and
spectroscopy studies suggest that the startburst started less than 4
Myr ago, while the low surface brightness regins are consistent with
an age below one Gyr \citep{2001AJ....121..169F}.

We suggest that new observational tests are needed to clarify the
kinematic nature of \tol\ and be sure about \tol's kinematic
structure. 
Integral field unit measurements resolving its spatial extent are up to the task. 
\tol\ spans a region of $4$ arcseconds,
an instrument such as the Multi Unit Spectroscopic Explorer (MUSE)
\citep{2014Msngr.157...13B} with its nominal $0.2$ arcseconds spatial
sampling over a $1.0$ arcminute field in wide-field mode could provide
a coarse mapping of different ionization lines to infer a kinematic map in the H$\alpha$ line,
following the kind of work presented by \citet{Herenz16} done with the Potsdam
Multi-Aperture Spectrometer (PMAS) \citep{PMAS} on nearby
\lya\ emitting galaxies.

Another observational test to confirm the picture provided by the
multiphase model includes the measurement of the \lya\ ionizing
continuum escape fraction. 
This model model predicts that averaging over all sightlines
it should be around $0.5^{+1.0}_{-0.4}$\%, with the possibility of strong
variations depending on viewing angle (i.e. some sightlines can have
an escape fraction close to $100\%$) \citep{Gronke2016} .

@@

It is the first time that an observed \lya\ profile can be fully
reproduced by either of these two kinematic conditions.
The structural conditions in \tol\ are similar to those in dwarf
ellipticals. 
However, it environment and star forming activity are radically
different.
We suggest that additional integral field unit observations are the
only way to be certain about the detailed kinematic structure in \tol.

@@

We acknowledge that a final confirmation of our kinematic findings has
to be provided by new observational tests of the kind we suggest.
If that were the case, this would open up a new line of more detailed
enquiries to understand how a compact dwarf galaxy can have the
structural characteristics similar to a dwarf elliptical while being
at the same time isolated and sustaining large starburst.
